---
title: "igraph: relationship analysis"
author: "Illarion  Jabine"
date: "30/12/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


### 1. Required packages:

In this manual I will work with igraph library.

* igraph: Network Analysis and Visualization. It can handle large graphs and provides functions for generating random
and regular graphs, graph visualization, centrality methods, etc.

```{r load libraries, message=FALSE, warning=FALSE}
library(igraph)

```


### 2. Key terms
* Assortativity
* Reciprocity
* Community detection
* Fast-greedy community detection
* Edge-betweenness community detection
Let's study how to partition networks into sub-networks and determine which vertices are more highly related to one another than others.


### 3. Useful Links & books
https://igraph.org/r/
https://en.wikipedia.org/wiki/Graph_(discrete_mathematics)
 
### 4. Creating a graph object

In this manual I will social network of people, see "1-creating-network" and "2-igraph-attributes-subsetting-deleting-edges-plot-weights". 
Edge list contains the names of people and time they were talking to each other.
Node list contains attributes such as age and sex.

First let's import into R edge and node lists.

```{r}

social_network_edges <- read_csv("social_network_edges.csv")
social_network_nodes <- read_csv("social_network_nodes.csv")
head(social_network_edges)
head(social_network_nodes)

# create an igraph object is by using graph_from_data_frame() function.
g1 <- graph_from_data_frame(d = social_network_edges, vertices = social_network_nodes, directed = FALSE)

```

### 5. Assortativity

Assortativity is a measure of how preferentially attached vertices are to other vertices with identical attributes. 
The degree of assortativity determines how preferentially attached are vertices to other vertices of a similar degree.
The assortativity coefficient measures the level of homophyly of the graph, based on some vertex labeling or values assigned to vertices. If the coefficient is high, that means that connected vertices tend to have the same labels or similar assigned values.

```{r}
# Convert the gender attribute into a numeric value
gender_values <- as.numeric(factor(V(g1)$gender))

# Calculate the assortativity of the network based on gender
assortativity(g1, gender_values)

# Calculate the assortativity degree of the network
assortativity.degree(g1, directed = FALSE)


```
Assortativity statistical tests

Using graph randomizations we can assess assortativity from statistical point of view. 
We will determine how likely the observed assortativity in the friendship network is given the genders of vertices by performing a randomization procedure. You will randomly permute the gender of vertices in the network 1000 times and recalculate the assortativity for each random network.

```{r}

# Calculate the assortativity observed in the network
obs_assort <- assortativity(g1, gender_values)

# Calculate the assortativity of the network randomizing the gender attribute 1000 times
# we use sample() to perform random samples and permutations with replace = FALSE
# table(sample(V(g1)$gender)) - proportion of F and M is the same all the time

# Initialize an empty list with 1000 elements
results <- vector('list', 1000)

# Run randomization process
for(i in 1:1000){
  results[[i]] <- assortativity(g1, sample(gender_values))
}

# Plot the distribution of assortativity values and add a red vertical line at the original observed value
hist(unlist(results))
abline(v = obs_assort, col = "red", lty = 3, lwd=3)

```

### 6. Reciprocity

Reciprocity of a directed network reflects the proportion of edges that are symmetrical, i.e.the proportion of outgoing edges that also have an incoming edge.
Example can be a network of people who send letters. Certain people may send letters to another but do not receive letters by that individual. 
Whereas other people may both send letters to each other and so would have a reciprocal tie.

Let's generate a random graph using sample_gnp() function according to the G(n,p) Erdos-Renyi model.
```{r}

g_recip <- sample_gnp(15, 5/15, directed=TRUE)

plot(g_recip,
      vertex.label.color = "black", 
      vertex.label.cex = 1.5,
      vertex.size = 1,
      edge.arrow.size = 0.3,
      edge.arrow.width = 0.5,
      edge.color = 'blue',
      main = "Reciprocity"
	  )

# Calculate the reciprocity of the graph
reciprocity(g_recip)

```
### 7. Community detection

Community detection in a network can be seen as clustering or grouping of nodes that share common attributes.
Community structure detection algorithms try to find dense subgraphs in directed or undirected graphs, by optimizing some criteria, and usually using heuristics.

7.1 Fast-greedy community detection

I will use Zachary Karate Club network, which contains 34 club members and 78 edges. Each edge indicates that those two club members interacted outside the karate club as well as at the club.
With Fast-greedy community detection we will determine how many sub-communities the network has and which club members belong to which subgroups.

```{r}
karate <- read_csv("karate.csv")

# create an igraph object is by using graph_from_data_frame() function.
g_karate <- graph_from_data_frame(d = karate, directed = FALSE)
```

```{r}
# Perform fast-greedy community detection on network graph
karate_comm <-  fastgreedy.community(g_karate)

# There are many S3 methods defined for object "communities" that igraph recognizes as being the output of community detection
## S3 method for class 'communities'
length(x)

# Determine sizes of each community
sizes(karate_comm)

# Determine which individuals belong to which community
membership(karate_comm)

algorithm(karate_comm)

merges(karate_comm)

crossing(karate_comm, graph)

code_len(karate_comm)

is_hierarchical(karate_comm)

# When plotting this object the network will be colored according to community membership.

plot(karate_comm, g_karate)

```

7.2 Edge-betweenness community detection

edge.betweenness.community() is another community detection method.
Community structure detection based on edge betweenness uses an idea that many networks consist of modules which are densely connected themselves but sparsely connected to other modules.
The idea of the edge betweenness based community structure detection is that it is likely that edges connecting separate modules have high edge betweenness as all the shortest paths from one module to another must traverse through them. So if we gradually remove the edge with the highest edge betweenness score we will get a hierarchical map, a rooted tree, called a dendrogram of the graph. The leafs of the tree are the individual vertices and the root of the tree represents the whole graph.

edge.betweeness.community() returns various information collected through the run of the algorithm. 

```{r}
# Perform edge-betweenness community detection on network graph
karate_comm_edg_betw = edge.betweenness.community(g_karate)

# Determine sizes of each community
sizes(karate_comm_edg_betw)
# Compared to Fast-greedy community detection this algorithm has detected 4 communities

```

Now let's compare two algoriths:
```{r}
# Plot community networks determined by fast-greedy and edge-betweenness methods side-by-side

par(mfrow = c(1, 2)) 
plot(karate_comm, g_karate)
plot(karate_comm_edg_betw, g_karate)
```
